---
title: "Medical Insurance: Who Becomes a High-Cost Patient?"
output:
  github_document:
    toc: true
    html_preview: false
    df_print: kable
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE, message = FALSE, fig.path = "figures/",
  fig.width = 7, fig.height = 4.2, dpi = 150
)
```

## 1. Goal

Medical costs are very different from person to person. An insurance company
wants to know **who is likely to be expensive**, so it can set fair prices and
plan for risk.

In this project we:

1. Label each person as **high cost** if their yearly medical charges are in the
   **top 30%**.
2. Build **logistic regression** models (a GLM with a binomial family) that
   predict the chance of being high cost from age, sex, BMI, number of
   children, smoking and region.
3. Compare 7 models with **AIC** and **10-fold cross-validation**, choose the
   best one and explain what it tells us.

## 2. Data

The data comes from Kaggle:
[Medical Insurance Price Prediction](https://www.kaggle.com/datasets/harishkumardatalab/medical-insurance-price-prediction).

| Column     | Meaning                                                  | Type     |
|------------|----------------------------------------------------------|----------|
| `age`      | Age of the person (years)                                | number   |
| `sex`      | `female` or `male`                                       | category |
| `bmi`      | Body Mass Index = weight (kg) / height (m)²              | number   |
| `children` | Number of children covered by the insurance              | number   |
| `smoker`   | `yes` or `no`                                            | category |
| `region`   | Home region in the US: `northeast`, `northwest`, `southeast`, `southwest` | category |
| `charges`  | Yearly medical costs billed by the insurance (USD)       | number   |

```{r libraries}
library(MASS)      # stepAIC()
library(bestglm)   # best subset selection
library(pROC)      # ROC curve and AUC
library(dplyr)     # loaded after MASS so that dplyr::select() is not masked
library(ggplot2)
library(scales)

theme_set(theme_minimal(base_size = 12))
smoker_colours    <- c(no = "#7FA7C9", yes = "#D1495B")
high_cost_colours <- c(No = "#A0A0A0", Yes = "#E08E2B")
```

```{r load-data}
raw <- read.csv("data/Medical_insurance.csv", stringsAsFactors = FALSE)
dim(raw)
head(raw)
```

## 3. Data cleaning

### 3.1 Missing values

```{r missing-values}
colSums(is.na(raw))
```

There are no missing values.

### 3.2 Duplicate rows

```{r duplicates}
sum(duplicated(raw))

# How many times does each different row appear in the file?
raw %>%
  count(across(everything()), name = "times_in_file") %>%
  count(times_in_file, name = "different_rows")
```

`r comma(sum(duplicated(raw)))` of the `r comma(nrow(raw))` rows
(`r percent(mean(duplicated(raw)), 1)`) are **exact copies** of another row.
Every column is the same, even `charges` with 5 decimals. Every different row
appears 2 or 4 times, so the file really has only
**`r comma(nrow(distinct(raw)))` different people**.

Why this is a problem:

- The model thinks it has twice as much data as it really has, so p-values
  look stronger than they are.
- In cross-validation, a copy of a test person can also be in the training
  data. The model has already "seen the answer", so test scores look better
  than they really are.

So we keep only one copy of each row.

```{r remove-duplicates}
ins <- distinct(raw)
nrow(ins)
```

### 3.3 Target variable: `high_cost`

```{r target}
cut_70 <- quantile(ins$charges, 0.70)
cut_70

ins <- ins %>%
  mutate(
    high_cost = factor(ifelse(charges >= cut_70, "Yes", "No"), levels = c("No", "Yes")),
    sex       = factor(sex),
    smoker    = factor(smoker),
    region    = factor(region)
  )

ins %>%
  count(high_cost) %>%
  mutate(percent = percent(n / sum(n), accuracy = 0.1))
```

A person is **high cost** if their charges are at least
**$`r comma(cut_70, accuracy = 0.01)`** (the 70th percentile).

- R uses the first level of each category as the **reference group**:
  `female`, `no` (non-smoker) and `northeast`.
- **Baseline:** if we always guess "No", we are right
  `r percent(mean(ins$high_cost == "No"), 0.1)` of the time. A useful model
  must beat this **70% baseline**.

## 4. Exploratory data analysis

### 4.1 High-cost rate by group

```{r rate-by-group}
rate_by <- function(var) {
  ins %>%
    group_by(group = as.character(.data[[var]])) %>%
    summarise(people = n(), high_cost = sum(high_cost == "Yes"), .groups = "drop") %>%
    mutate(variable = var, high_cost_rate = percent(high_cost / people, 0.1)) %>%
    select(variable, group, people, high_cost, high_cost_rate)
}

bind_rows(rate_by("smoker"), rate_by("sex"), rate_by("region"))
```

- **Smoking** makes the biggest difference: almost every smoker is high cost.
- **Sex** makes only a small difference.
- **Region:** the southwest has the lowest rate and the southeast the highest.

### 4.2 Smoking status by cost group

```{r plot-smoker-by-cost}
ggplot(ins, aes(x = high_cost, fill = smoker)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = label_percent()) +
  scale_fill_manual(values = smoker_colours) +
  labs(title = "Smoking status by cost group",
       x = "High cost?", y = "Share of people", fill = "Smoker")
```

In the low-cost group almost nobody smokes
(`r percent(mean(ins$smoker[ins$high_cost == "No"] == "yes"), 0.1)`).
In the high-cost group most people smoke
(`r percent(mean(ins$smoker[ins$high_cost == "Yes"] == "yes"), 0.1)`).

### 4.3 Sex by cost group

```{r plot-sex-by-cost}
ggplot(ins, aes(x = high_cost, fill = sex)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = label_percent()) +
  labs(title = "Sex by cost group",
       x = "High cost?", y = "Share of people", fill = "Sex")
```

The male/female mix is almost the same in both groups (men:
`r percent(mean(ins$sex[ins$high_cost == "No"] == "male"), 0.1)` of the
low-cost group and
`r percent(mean(ins$sex[ins$high_cost == "Yes"] == "male"), 0.1)` of the
high-cost group). Sex does not look important.

### 4.4 Smoking status by region

```{r plot-smoker-by-region}
ggplot(ins, aes(x = region, fill = smoker)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = label_percent()) +
  scale_fill_manual(values = smoker_colours) +
  labs(title = "Smoking status by region",
       x = "Region", y = "Share of people", fill = "Smoker")
```

```{r smoking-rate-by-region}
ins %>%
  group_by(region) %>%
  summarise(smokers = percent(mean(smoker == "yes"), 0.1))
```

Smoking rates are similar in all regions. The southeast has a few more
smokers, which partly explains its higher high-cost rate.

### 4.5 Charges by smoking status

```{r plot-charges-by-smoker}
ggplot(ins, aes(x = smoker, y = charges, fill = smoker)) +
  geom_boxplot(alpha = 0.8) +
  scale_y_continuous(labels = label_dollar()) +
  scale_fill_manual(values = smoker_colours) +
  labs(title = "Charges by smoking status", x = "Smoker", y = "Charges (USD)") +
  theme(legend.position = "none")
```

The median charge of smokers is
$`r comma(median(ins$charges[ins$smoker == "yes"]))`, about
`r round(median(ins$charges[ins$smoker == "yes"]) / median(ins$charges[ins$smoker == "no"]), 1)`
times the median of non-smokers
($`r comma(median(ins$charges[ins$smoker == "no"]))`). Smokers' charges are
also much more spread out.

### 4.6 Charges by cost group

```{r plot-charges-by-cost}
ggplot(ins, aes(x = high_cost, y = charges, fill = high_cost)) +
  geom_boxplot(alpha = 0.8) +
  geom_hline(yintercept = cut_70, linetype = "dashed") +
  scale_y_continuous(labels = label_dollar()) +
  scale_fill_manual(values = high_cost_colours) +
  labs(title = "Charges by cost group (dashed line = cutoff)",
       x = "High cost?", y = "Charges (USD)") +
  theme(legend.position = "none")
```

This checks that the target was made correctly: every "No" is below the
cutoff and every "Yes" is above it. The high-cost group is also much more
spread out (from $`r comma(min(ins$charges[ins$high_cost == "Yes"]))` to
$`r comma(max(ins$charges[ins$high_cost == "Yes"]))`).

### 4.7 BMI by sex

```{r plot-bmi-by-sex}
ggplot(ins, aes(x = sex, y = bmi, fill = sex)) +
  geom_boxplot(alpha = 0.8) +
  labs(title = "BMI by sex", x = "Sex", y = "BMI") +
  theme(legend.position = "none")
```

BMI looks almost the same for women (median
`r round(median(ins$bmi[ins$sex == "female"]), 1)`) and men (median
`r round(median(ins$bmi[ins$sex == "male"]), 1)`). Both groups have some very
high BMI values.

### 4.8 Charges by age and smoking status

```{r plot-charges-by-age}
ggplot(ins, aes(x = age, y = charges, colour = smoker)) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = cut_70, linetype = "dashed") +
  annotate("text", x = 36, y = cut_70, label = "high-cost cutoff",
           hjust = 0, vjust = 1.5, size = 3.5) +
  scale_y_continuous(labels = label_dollar()) +
  scale_colour_manual(values = smoker_colours) +
  labs(title = "Charges by age and smoking status",
       x = "Age", y = "Charges (USD)", colour = "Smoker")
```

This plot explains most of the project:

- Almost all smokers are above the cutoff line.
- For most non-smokers, charges grow slowly with age and only reach the
  cutoff line at around age 60.
- A small group of non-smokers is expensive at any age. Nothing in the data
  explains why.

So the link between age and high cost is **not a straight line**: for
non-smokers, the chance stays low for a long time and then jumps up at older
ages. This is why we also try curved terms (age², age³) in the models.

## 5. Logistic regression models

Logistic regression models the probability $p$ that a person is high cost:

$$\log\left(\frac{p}{1-p}\right) = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \dots$$

- A positive coefficient means a higher chance of high cost.
- $e^{\beta}$ is the **odds ratio**: how many times the odds change when $x$
  goes up by 1.

We compare models with **AIC** (Akaike Information Criterion):
AIC = −2 × log-likelihood + 2 × (number of parameters).
A lower AIC means a better balance between fitting the data well and keeping
the model simple.

| Model | Idea | Terms |
|---|---|---|
| M1 Full | Main effects + BMI² + interactions with smoking | age, bmi, bmi², children, sex, smoker, region, smoker × bmi, smoker × age |
| M2 Reduced | M1 without the interactions | age, bmi, bmi², children, sex, smoker, region |
| M3 Quadratic | M2 + age² | M2 + age² |
| M4 Cubic | M3 + age³ + bmi³ | M3 + age³, bmi³ |
| M5 Stepwise AIC | Backward selection starting from M4 | chosen by AIC |
| M6 bestglm (AIC) | Try every subset of the main effects | chosen by AIC |
| M7 bestglm (CV) | Try every subset of the main effects | chosen by 10-fold CV |

### 5.1 M1: Full model

```{r m1-full}
m1_full <- glm(
  high_cost ~ age + bmi + I(bmi^2) + children + sex + smoker + region +
    smoker:bmi + smoker:age,
  data = ins, family = binomial
)

m1_full$converged
round(coef(summary(m1_full)), 3)
```

**M1 does not work.** R warns that the algorithm did not converge, and the
smoker terms have huge coefficients (`smokeryes` =
`r comma(coef(m1_full)[["smokeryes"]])`) with huge standard errors. This
problem is called **separation**. Here is the reason:

```{r low-cost-smokers}
ins %>% filter(smoker == "yes", high_cost == "No")
```

Only `r sum(ins$smoker == "yes" & ins$high_cost == "No")` of the
`r sum(ins$smoker == "yes")` smokers are **not** high cost, and both are
18-year-old men with a low BMI. With the interaction terms, smokers get their
own age and BMI slopes, so the model can draw a line that separates these 2
people perfectly from the other smokers. To make this "perfect", the
coefficients keep growing and the fit never finishes.

So the coefficients, p-values and AIC of M1 **cannot be trusted**, and M1 cannot
be the final model. It also tells us that this data is too small to estimate
smoker × age or smoker × BMI effects for a yes/no high-cost target.

### 5.2 M2, M3, M4: Reduced, quadratic and cubic models

```{r m2-m4}
m2_reduced <- glm(
  high_cost ~ age + bmi + I(bmi^2) + children + sex + smoker + region,
  data = ins, family = binomial
)

m3_quadratic <- glm(
  high_cost ~ age + I(age^2) + bmi + I(bmi^2) + children + sex + smoker + region,
  data = ins, family = binomial
)

m4_cubic <- glm(
  high_cost ~ age + I(age^2) + I(age^3) + bmi + I(bmi^2) + I(bmi^3) +
    children + sex + smoker + region,
  data = ins, family = binomial
)

c(m2 = m2_reduced$converged, m3 = m3_quadratic$converged, m4 = m4_cubic$converged)
AIC(m2_reduced, m3_quadratic, m4_cubic)
```

- Without the interactions, the separation problem is gone: all three models
  converge.
- Adding age² (M3) lowers the AIC a lot, and adding age³ and bmi³ (M4) lowers
  it again. This matches the plot in 4.8: the effect of age is curved.

### 5.3 M5: Stepwise AIC

Backward stepwise selection starts from a big model and removes one term at a
time, as long as the AIC goes down. We start from **M4**, the largest model
that converges. (Starting from M1 is not useful, because its AIC is not
reliable.)

```{r m5-stepwise}
m5_stepwise <- stepAIC(m4_cubic, direction = "backward", trace = FALSE)
m5_stepwise$anova
formula(m5_stepwise)
```

Stepwise removes only `bmi³`. Removing any other term would not lower the AIC.

### 5.4 M6 and M7: best subset selection with `bestglm`

`bestglm` fits **every possible subset** of the main effects
(2⁸ = 256 models) and picks the best one. It needs a data frame with only
numbers and the outcome `y` in the last column, so we turn each category into
0/1 columns. Region gets 3 columns because `northeast` is the reference group.

```{r bestglm-data}
ins_best <- ins %>%
  transmute(
    age, bmi, children,
    sex_male   = as.integer(sex == "male"),
    smoker_yes = as.integer(smoker == "yes"),
    reg_nw     = as.integer(region == "northwest"),
    reg_se     = as.integer(region == "southeast"),
    reg_sw     = as.integer(region == "southwest"),
    y          = as.integer(high_cost == "Yes")
  )
```

**M6: best subset by AIC**

```{r m6-bestglm-aic}
best_aic <- bestglm(ins_best, family = binomial, IC = "AIC",
                    method = "exhaustive", TopModels = 5)
best_aic$BestModels

# Refit the chosen subset as a normal glm on ins_best (needed for cross-validation later)
m6_best_aic <- glm(formula(best_aic$BestModel), data = ins_best, family = binomial)
print(formula(m6_best_aic), showEnv = FALSE)
```

The table shows the top 5 subsets. (bestglm's `Criterion` does not count the
intercept, so it is exactly 2 lower than R's `AIC()`.)

**M7: best subset by 10-fold cross-validation**

```{r m7-bestglm-cv}
set.seed(999)
best_cv <- bestglm(ins_best, family = binomial, IC = "CV",
                   CVArgs = list(Method = "HTF", K = 10, REP = 1),
                   method = "exhaustive")
m7_best_cv <- glm(formula(best_cv$BestModel), data = ins_best, family = binomial)
print(formula(m7_best_cv), showEnv = FALSE)
```

The HTF method uses the **one-standard-error rule**: it picks the simplest
model whose CV error is close to the best one. That is why M7 keeps only age
and smoker.

## 6. Model comparison

AIC measures the fit on the data that the model was trained on. To check how
well each model predicts **new people**, we use **10-fold cross-validation
(CV)**:

1. Split the `r comma(nrow(ins))` people into 10 random groups (folds).
2. Train the model on 9 folds and predict the 10th fold.
3. Repeat 10 times, so every person gets one prediction from a model that
   never saw them.

All models use the **same folds**, so the comparison is fair. We then compute:

- **Accuracy:** share of correct yes/no predictions (predict "Yes" when the
  probability is 0.5 or more).
- **AUC:** the chance that the model gives a random high-cost person a higher
  probability than a random low-cost person (1 = perfect, 0.5 = coin flip).

```{r cross-validation}
models <- list(
  "M1 Full (interactions)" = m1_full,
  "M2 Reduced"             = m2_reduced,
  "M3 Quadratic"           = m3_quadratic,
  "M4 Cubic"               = m4_cubic,
  "M5 Stepwise AIC"        = m5_stepwise,
  "M6 bestglm (AIC)"       = m6_best_aic,
  "M7 bestglm (CV)"        = m7_best_cv
)

set.seed(999)
folds  <- sample(rep(1:10, length.out = nrow(ins)))
actual <- as.integer(ins$high_cost == "Yes")

# Out-of-fold probability for every person
cv_predict <- function(model) {
  data <- model$data
  pred <- numeric(nrow(data))
  for (k in 1:10) {
    test <- folds == k
    fit  <- suppressWarnings(glm(formula(model), data = data[!test, ], family = binomial))
    pred[test] <- predict(fit, newdata = data[test, ], type = "response")
  }
  pred
}

cv_pred <- lapply(models, cv_predict)

comparison <- data.frame(
  model       = names(models),
  parameters  = sapply(models, function(m) length(coef(m))),
  converged   = sapply(models, function(m) m$converged),
  AIC         = round(sapply(models, AIC), 1),
  cv_accuracy = percent(sapply(cv_pred, function(p) mean((p >= 0.5) == actual)), 0.1),
  cv_auc      = round(sapply(cv_pred, function(p) {
    as.numeric(auc(actual, p, levels = c(0, 1), direction = "<", quiet = TRUE))
  }), 3),
  nonsmokers_predicted_high = sapply(cv_pred, function(p) sum(p >= 0.5 & ins$smoker == "no")),
  row.names   = NULL
)
comparison
```

**Two simple rules to compare with:**

- Always guess "No": `r percent(mean(actual == 0), 0.1)` accuracy.
- Guess "Yes" only for smokers:
  `r percent(mean((ins$smoker == "yes") == (actual == 1)), 0.1)` accuracy.

What we learn:

- **M1, M2, M6 and M7 are only as good as the smoker rule** (90.2–90.3%
  accuracy). With a straight-line age effect, they (almost) never predict
  "high cost" for a non-smoker (last column), so in practice they just say
  "Yes" for smokers (see 7.2).
- **The curved-age models (M3, M4, M5) are clearly better** (92.4–92.5%
  accuracy and a higher AUC). The curve lets them also flag older
  non-smokers. The differences among these three are only 1 or 2 people.
- **M5 has the lowest AIC** of all models that converge, and its CV scores are
  almost the same as the best ones.

**Final model: M5 (Stepwise AIC).**

## 7. Final model

```{r final-summary}
final_model <- m5_stepwise
summary(final_model)
```

### 7.1 Odds ratios

```{r odds-ratios}
or_table <- exp(cbind(odds_ratio = coef(final_model), confint(final_model)))
round(or_table[c("children", "sexmale", "smokeryes",
                 "regionnorthwest", "regionsoutheast", "regionsouthwest"), ], 2)
```

An odds ratio above 1 means higher odds of high cost; below 1 means lower odds.
If the 95% interval includes 1, the effect is not clear.

- **Smoker:** the odds are about `r comma(or_table["smokeryes", 1])` times
  higher. This number is extreme because
  `r sum(ins$smoker == "yes" & ins$high_cost == "Yes")` of
  `r sum(ins$smoker == "yes")` smokers are high cost, so the exact size is very
  uncertain (wide interval). But it is clearly the strongest factor.
- **Children:** each extra child multiplies the odds by
  `r round(or_table["children", 1], 2)`
  (+`r percent(or_table["children", 1] - 1, 1)`).
- **Region:** people in the southwest have
  `r percent(1 - or_table["regionsouthwest", 1], 1)` lower odds than people in
  the northeast. Northwest and southeast are not clearly different.
- **Sex:** the interval includes 1, so there is no clear difference between men
  and women.
- **Age and BMI** are curved (several terms work together), so single odds
  ratios are hard to read. We show age in a plot below. For BMI, the model
  says the chance goes up until a BMI of about
  `r round(-coef(final_model)[["bmi"]] / (2 * coef(final_model)[["I(bmi^2)"]]), 1)`
  and then goes down a little. This effect is much smaller than smoking or
  age.

### 7.2 Effect of age for non-smokers

For smokers, the chance of high cost is almost 100% at every age, so age
matters mostly for **non-smokers**. The points below show the real share of
high-cost non-smokers at each age. The lines show the average predicted
probability (models fitted on all the data) of a model with a straight-line
age effect (M2) and of the final model (M5).

```{r plot-age-effect}
nonsmokers <- ins %>%
  mutate(p_linear = fitted(m2_reduced), p_final = fitted(final_model)) %>%
  filter(smoker == "no")

age_effect <- nonsmokers %>%
  group_by(age) %>%
  summarise(observed = mean(high_cost == "Yes"),
            p_linear = mean(p_linear),
            p_final  = mean(p_final))

ggplot(age_effect, aes(x = age)) +
  geom_point(aes(y = observed), colour = "grey40") +
  geom_line(aes(y = p_linear, colour = "M2: straight-line age"), linewidth = 1) +
  geom_line(aes(y = p_final, colour = "M5 (final): curved age"), linewidth = 1) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  scale_y_continuous(labels = label_percent()) +
  scale_colour_manual(values = c("#7FA7C9", "#D1495B")) +
  labs(title = "Non-smokers: chance of high cost by age",
       subtitle = "Points = real share at each age. Lines = average model prediction.",
       x = "Age", y = "Chance of high cost", colour = NULL) +
  theme(legend.position = "bottom")
```

```{r nonsmokers-predicted-high}
nonsmokers %>%
  summarise(
    highest_chance_M2 = percent(max(p_linear), 0.1),
    predicted_high_M2 = sum(p_linear >= 0.5),
    highest_chance_M5 = percent(max(p_final), 0.1),
    predicted_high_M5 = sum(p_final >= 0.5),
    youngest_predicted_high_M5 = min(age[p_final >= 0.5])
  )
```

- For non-smokers, the real chance of high cost stays low (about 5–15%) until
  the late 50s, then rises fast after 60.
- The straight-line model (M2) cannot bend like this. Its highest chance for
  any non-smoker is only `r percent(max(nonsmokers$p_linear), 0.1)`, so it
  says "high cost" for only `r sum(nonsmokers$p_linear >= 0.5)` of the
  `r comma(nrow(nonsmokers))` non-smokers.
- The final model (M5) follows the jump at older ages, so it flags
  `r sum(nonsmokers$p_final >= 0.5)` non-smokers, all aged
  `r min(nonsmokers$age[nonsmokers$p_final >= 0.5])` or older. This is where
  its extra accuracy comes from.

### 7.3 Performance on new data (10-fold CV)

All numbers here come from the out-of-fold predictions in section 6, so the
model never saw the people it is scored on.

```{r plot-roc, fig.width = 5.5, fig.height = 5.5}
final_pred <- cv_pred[["M5 Stepwise AIC"]]
roc_final  <- roc(actual, final_pred, levels = c(0, 1), direction = "<", quiet = TRUE)
ci.auc(roc_final)

ggroc(roc_final, legacy.axes = TRUE, colour = "#D1495B", linewidth = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  coord_equal() +
  labs(title = sprintf("Final model ROC (10-fold CV), AUC = %.3f", auc(roc_final)),
       x = "False positive rate (1 - specificity)",
       y = "True positive rate (sensitivity)")
```

```{r confusion-matrix}
predicted <- factor(ifelse(final_pred >= 0.5, "Yes", "No"), levels = c("No", "Yes"))
confusion <- table(Predicted = predicted, Actual = ins$high_cost)
confusion

tp <- confusion["Yes", "Yes"]; tn <- confusion["No", "No"]
fp <- confusion["Yes", "No"];  fn <- confusion["No", "Yes"]

data.frame(
  metric = c("Accuracy",
             "Sensitivity (share of high-cost people we find)",
             "Specificity (share of low-cost people we find)",
             "Precision (share of 'Yes' predictions that are right)"),
  value  = percent(c((tp + tn) / sum(confusion), tp / (tp + fn),
                     tn / (tn + fp), tp / (tp + fp)), 0.1)
)
```

Who are the high-cost people that the model misses?

```{r missed-people}
missed <- ins %>% filter(predicted == "No", high_cost == "Yes")
missed %>% count(smoker, age_group = ifelse(age >= 60, "60+", "under 60"))
```

- The model is right for `r percent((tp + tn) / sum(confusion), 0.1)` of people
  (baseline: 70%) and almost never calls a low-cost person "high cost"
  (only `r fp` of `r tn + fp`).
- Of the `r fn` high-cost people it misses,
  `r sum(missed$smoker == "no" & missed$age < 60)` are non-smokers under 60.
  They are the small group from 4.8 that is expensive for reasons that are
  **not in the data**, so they are very hard to find with these 6 columns.

## 8. Key findings

1. **Smoking is by far the strongest factor.**
   `r sum(ins$smoker == "yes" & ins$high_cost == "Yes")` of
   `r sum(ins$smoker == "yes")` smokers are high cost, compared with
   `r percent(mean(ins$high_cost[ins$smoker == "no"] == "Yes"), 0.1)` of
   non-smokers.
2. **Age matters in a curved way.** For non-smokers, the risk is low until
   the late 50s and rises fast after 60. Models with age² and age³ beat the
   straight-line models.
3. **More children means a higher risk** (odds × `r round(or_table["children", 1], 2)`
   per child), and the **southwest has a lower risk** than the northeast.
4. **Sex does not have a clear effect**, and BMI has only a small effect on
   this yes/no target.
5. The **final model (M5, stepwise AIC)** has the lowest AIC among the models
   that converge and a CV accuracy of `r percent((tp + tn) / sum(confusion), 0.1)`
   with an AUC of `r round(as.numeric(auc(roc_final)), 3)`.
6. The full model with smoker interactions **cannot be fitted** (separation),
   because only 2 smokers are not high cost.

## 9. Limitations and next steps

- **A yes/no target loses information.** For example, BMI changes smokers'
  charges a lot (median
  $`r comma(median(ins$charges[ins$smoker == "yes" & ins$bmi >= 30]))` with a
  BMI of 30 or more vs
  $`r comma(median(ins$charges[ins$smoker == "yes" & ins$bmi < 30]))` below 30),
  but almost all smokers are already above the cutoff, so the model cannot see
  this. *Next step:* model `charges` directly, for example with a Gamma GLM
  with a log link.
- **Separation.** Smoking almost perfectly predicts high cost, so smoker
  interactions cannot be estimated and the smoker odds ratio is very uncertain.
  *Next step:* Firth's penalized logistic regression (`logistf`) or LASSO
  (`glmnet`).
- **Variable selection uses the same data.** Stepwise and best subset
  selection make p-values look better than they are. In our CV, the formula
  was chosen once on all data (the selection was not repeated inside each
  fold), so the CV scores may be a little optimistic. *Next step:* nested CV.
- **Missing information.** Only 6 predictors, and some non-smokers are
  expensive at any age for reasons that are not in the data.
- **No external test data.** All results come from one dataset of
  `r comma(nrow(ins))` people.

## 10. Session info

```{r session-info}
sessionInfo()
```
